Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-08 arxiv/diff_cov.tex
section 4 polish

Our key result characterizes the asymptotic limit of the sum of two arbitrarily different sample covariance matrices.
To motivate our study, consider a special case where $A$ is equal to one.
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align}\label{def hatsig}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where } \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariances, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.
The empirical loss of $\hat{\beta}_2^{\MTL}$ is close to its expectation similar to Claim \ref{claim_pred_err}.
Hence, we consider the expected loss conditional on the covariates as follows:
\begin{align}\label{eq_diff_cov}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	\bignorm{(\Sigma^{(2)})^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2
	+ \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}.
\end{align}

@@

\begin{theorem}\label{thm_main_RMT}
	Suppose that $X^{(1)}$ and $X^{(2)}$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	Assume that: a) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$; b) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have the following limits:

	\noindent (i) The variance equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following two equations:
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\quad a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
	\noindent (ii) For any fixed vector $w\in\real^p$ with unit norm, the bias equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following two equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_1 a_4 - a_2 a_3)}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
\end{theorem}
Our result extends a well-known result on the inverse of a sample covariance matrix.
In particular, when $X_1$ is zero, we obtain Fact \ref{lem_minv}(ii) as a corollary.
To see this, we solve equation \eqref{eq_a12extra} and obtain $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply these solutions to the variance limit.
One can see that both the variance limit and the bias limit depend heavily on both tasks' samples sizes and the covariate shift matrix $M$.
Next, we illustrate how varying sample sizes and covariate shifts affects hard parameter sharing, respectively.

@@

	In applying Theorem \ref{thm_main_RMT} to the above setting, we first solve the self-consistent equations \eqref{eq_a12extra} by using $\lambda_i = 1$ for all $1\le i\le p$.
	This gives us the variance limit $\tr[(a_1\Sigma^{(1)} + a_2\Sigma^{(2)})^{-1}] / (n_1 + n_2) = \frac{p}{n_1 + n_2 - p}$, since we have that $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2 - p}$.
	Secondly, we solve the self-consistent equations \eqref{eq_a34extra} given that we already know $a_1$ and $a_2$.
	This gives us the bias limit.
	Combined together, we obtain the following corollary, which requires natural assumptions over the random-effects model.

\begin{corollary}\label{cor_MTL_loss}
	In the setting of Example \ref{ex_sample_ratio}, assume that
	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
	(i) task two's sample size is at least $n_2 \ge \frac{3}{2}p$;
	(ii) the noise variance is smaller than the signal strength: $\sigma^2 \lesssim  \kappa^2$;
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	%\be\label{choiceofpara0}
%	(ii) the task-specific variance of $\beta_i$ is much smaller than the signal strength {\color{red}$d^2 = \oo( {\kappa^2})$}; \HZ{what does $\ll$ mean exactly?}
%	(iii) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$.
	(iii) the task-specific variance of $\beta_i$ is much smaller than the shared signal strength: $d^2 \le p^{-c_{\varphi}}{\kappa^2}$.
	Let $\varepsilon = (1 + \sqrt{p/n_1})^ 4 - 1$, which decreases as $n_1$ increases.
	The prediction loss of hard parameter sharing for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	  \left|L(\hat{\beta}_2^{\MTL}) - \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right| \le \varepsilon \cdot \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} + p^{-c_{\varphi}} \cdot \OO(d^2 + \sigma^2).
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
%	 {\color{red}[FY: the error also contains $p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) $, both of which cannot be omitted, because (i) there is no assumption on the upper bound of $\kappa^2$, and (ii) we do not necessarily have $c_\varphi<1/4$. We can decide how to present the result concisely (for instance we can impose an upper bound on $\kappa^2$ and that $c_\varphi<1/4$), but it needs to be correct.]}
	 \end{corollary}

@@

Our result shows that the prediction loss of hard parameter sharing consists of a bias term that scales with task-specific variance $d^2$ and a noise variance term that scales with $\sigma^2$.
In the following, we explain an interesting phenomenon, where adding task one's samples helps task two initially, but hurts eventually.
Consider the prediction loss if we fix $n_2$ and increase $n_1$.
We show that as a function of $n_1$, the limit equation always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ with $n_1$ going to infinity.
We describe an informal proof.
First, the noise variance term in Corollary \ref{cor_MTL_loss} always reduces as $n_1$ increases.
Second, the bias term always increases as $n_1$ increases, which can be verified by calculating the derivative of the bias term.
Moreover, near the point where $n_1$ is zero, one can verify that bias increases less than the decrease in variance, and there is only one positive root for the derivative being zero.
%Combined together, we observe that the \textit{optimal} sample ratio between $n_1$ and $n_2$ should be among one of following three cases:
%(i) $n_1 = 0$, in which case the prediction loss reduces to $\frac{\sigma^2 p}{n_2 - p}$;
%(ii) $n_1 = \infty$, in which case the prediction loss reduces to $d^2$;
%(iii) the critial point where the derivative of the bias and the variance terms w.r.t. $n_1$ equals zero.
The proof of Corollary \ref{cor_MTL_loss} can be found in Appendix \ref{app_iso_cov}.

%Based on this observation, we have the following trend when we fix $n_2$ and increase $n_1$:
%\begin{itemize}
%	\item If $d^2$ is smaller than $\frac{\sigma^2 p}{n_2 - p}$, then the prediction loss of hard parameter sharing is always lower than STL. That is, task one always transfers positively to task two.
%	\item Otherwise, task one transfers positively to task two initially, but transfers negatively eventually.
%\end{itemize}
%Therefore, while the variance of MTL still reduces compared to STL,
%If the bias increases more than the amount of variance reduced, we observe negative transfer.
%of MTL increases and converges to $pd^2$ as the sample ratio $\rho_1/\rho_2$ increases, and the variance reduction $\frac{\sigma^2}{ \rho_2-1}  -\frac{\sigma^2}{\rho_1+\rho_2-1}$ also increases and converges to $\frac{\sigma^2}{\rho_2-1}$ as $\rho_1/\rho_2$ increases. The tradeoff between these two terms depends on $\rho_1$ in a nonlinear way that is more complicated than the one in the case of varying model distance. But from our results, one can check the following phenomena. If $\frac{\sigma^2}{\rho_2-1} > pd^2$, then the transfer is always positive. On the other hand, if $\frac{\sigma^2}{\rho_2-1} < pd^2$, then there is a transition from the positive transfer to negative transfer as $\rho_1/\rho_2$ increases.
%Therefore, while the variance of MTL still reduces compared to STL,
%If the bias increases more than the amount of variance reduced, we will observe negative transfer.
%Figure \ref{fig_size} provides a simulation result for such a setting (described in Appendix \ref{app_synthetic}).
%We observe that as $n_1 / n_2$ increases, there is a transition from positive to negative transfer.
%There are several unexplained observations near $y = 0$ caused by the concentration error on the right-hand side of \eqref{cor_MTL_error}.
\end{example}

\begin{example}[Covariate shift]\label{ex_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.
Our second example studies the impact of varying covariate shifts on the variance limit.
%As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
Recall that the matrix $M$ captures the geometry of covariate shifts.
Suppose that half of the singular values of $M$ are equal to a parameter $\lambda > 0$ and the other half are equal to $\lambda^{-1}$.
We observe the following dichotomy on the impact of covariate shifts.
\begin{itemize}
	\item If $n_1 \ge n_2$, then the variance (limit) of hard parameter sharing is smallest when $\lambda = 1$ or there is no covariate shift.
	\item If $n_1 < n_2$, then having covariate shifts ($\lambda \neq 1$) always helps reduce variance (limit).
\end{itemize}
To see this, we first write the variance limit for different values of $\lambda$ as follows
\begin{align*}
	\bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}
	= \bigtr{(a_1 M^{\top} M + a_2 \id)^{-1}}
	= \frac{1}{2}\bigbrace{(\lambda^{-2} a_1 + a_2)^{-1} + (\lambda^2 a_1 + a_2)^{-1}}.
\end{align*}
Denote the above function by $f(\lambda)$.
We compare $f(\lambda)$ with $f(1)$, and observe that
\[ 2f(\lambda) - 2f(1) = {(\lambda^2 a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1})^{-1}
	+ (\lambda^{-2} a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1)^{-1}  - \frac{2(n_1 + n_2)}{n_1 + n_2 - p}, \]
where we used the first part of equation \eqref{lem_cov_shift_eq}.
Next, one can verify that the above equation is equal to $(2a_1 - \frac{n_1 + n_2 - p} {n_1 + n_2}) g(\lambda, a_1)$, for a fixed and positive function $g(\lambda, a_1)$ (details omitted).
Hence, whether or not $f(\lambda)$ is small than $f(1)$ is decided by $a_1$.
To solve $a_1$, we present an informal argument that $a_1 \ge \frac{(n_1 + n_2 - p)}{2(n_1 + n_2)}$ if and only if $n_1 \ge n_2$.
Intuitively, with the current spectrum of $M$, $a_1$ and $a_2$ have a symmetric role in the self-consistent equations.
Hence, when $n_1 \ge n_2$, we expect that $a_1 \ge a_2$, hence $a_1 \ge \frac{1}{2}(1 - \frac{p}{n_1 + n_2})$.
Vice serva when $n_1 < n_2$.
A formal proof requires solving the self-consistent equations \eqref{lem_cov_shift_eq} by replacing half of the singular with $\lambda$ and the other half with $\lambda^{-1}$.
We omit the details.
Thus, we conclude that if $n_1 \ge n_2$, then $f(\lambda) > f(1)$; if $n_1< n_2$, then $f(\lambda)< f(1)$.
\end{example}
%\begin{proposition}[Source/target sample ratio]\label{prop_data_size}
%	In the isotropic model, suppose that $\rho_1 > 40$ and $\rho_2 > 110$ are fixed constants, and $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
%	Then we have that
%	\begin{itemize}
%		\item \textbf{Positive transfer:} If $\frac{n_1}{n_2} = \frac{\rho_1}{\rho_2} < \frac{1}{\nu} \cdot \frac{1 - 2\rho_2^{-1}}{\Psi(\beta_1, \beta_2) (\rho_2 - 1) - \nu^{-1}}$, then w.h.p. $$\te(\hat{\beta}_2^{\MTL}) < \te(\hat{\beta}_2^{\STL}).$$
%		\item \textbf{Negative transfer:} If $\frac{n_1}{n_2} = \frac{\rho_1}{\rho_2} > {\nu} \cdot \frac{1 - 2\rho_2^{-1}}{\Psi(\beta_1, \beta_2) (\rho_2 - 3/2) - \nu}$, then w.h.p. $$\te(\hat{\beta}_2^{\MTL}) > \te(\hat{\beta}_2^{\STL}).$$
%	\end{itemize}
%\end{proposition}
%Proposition \ref{prop_data_size} describes the bias-variance tradeoff in terms of the sample ratio $\rho_1 / \rho_2$.

>>>>

Our key result characterizes the asymptotic limit of the sum of two arbitrarily different sample covariance matrices.
To motivate our study, consider a special case where $A$ is equal to one.
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align*}%\label{def hatsig}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where } \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align*}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariances, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.
The empirical loss of $\hat{\beta}_2^{\MTL}$ is close to its expectation similar to Claim \ref{claim_pred_err}.
Hence, we consider the expected loss conditional on the covariates as follows:
\begin{align}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	& \bignorm{(\Sigma^{(2)})^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}

@@

\begin{theorem}\label{thm_main_RMT}
	Suppose that $X^{(1)} \in \real^{n_1 \times p}$ and $X^{(2)} \in \real^{n_2 \times p}$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	Assume that: a) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$; b) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have the following limits:

	\noindent (i) The variance equation \eqref{eq_variance_2task} satisfies the following limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following self-consistent equations
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\quad a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
	\noindent (ii) The bias equation \eqref{eq_bias_2task} satisfies the following limit for any fixed vector $w\in\real^p$ with unit norm,
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following self-consistent equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_1 a_4 - a_2 a_3)}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
\end{theorem}
Our result extends a well-known result on the inverse of a sample covariance matrix.
In particular, when $X_1$ is zero, we obtain Fact \ref{lem_minv}(ii) as a corollary.
To see this, we solve equation \eqref{eq_a12extra} and obtain $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply these solutions to the variance limit.
One can see that both the variance limit and the bias limit depend heavily on both tasks' samples sizes and the covariate shift matrix $M$.
Next, we illustrate how varying sample sizes and covariate shifts affects hard parameter sharing, respectively.

@@

	Applying Theorem \ref{thm_main_RMT} to the above setting, we first solve the self-consistent equations \eqref{eq_a12extra} by using $\lambda_i = 1$ for all $1\le i\le p$.
	This gives us the variance limit
	\[ \frac{1}{n_1 + n_2} \tr[\Sigma^{(2)} (a_1\Sigma^{(1)} + a_2\Sigma^{(2)})^{-1}] = \frac{p}{(n_1 + n_2)(a_1 + a_2)} = \frac{p}{n_1 + n_2 - p}, \]
	since $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2 - p}$.
	Similarly, for the bias limit, we solve the self-consistent equations \eqref{eq_a34extra} since we have obtained $a_1, a_2$.
	Combined together, we obtain the following corollary of Theorem \ref{thm_main_RMT}.
\end{example}

\begin{corollary}\label{cor_MTL_loss}
	In the setting of Example \ref{ex_sample_ratio}, assume that
	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
	(i) both tasks sample sizes are at least $\frac{3}{2}p$;
	(ii) noise variance is smaller than the signal strength: $\sigma^2 \lesssim  \kappa^2$;
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	%\be\label{choiceofpara0}
%	(ii) the task-specific variance of $\beta_i$ is much smaller than the signal strength {\color{red}$d^2 = \oo( {\kappa^2})$}; \HZ{what does $\ll$ mean exactly?}
%	(iii) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$.
	(iii) task-specific variance is much smaller than the shared signal strength: $d^2 \le p^{-c_{\varphi}}{\kappa^2}$.
	Let $\varepsilon = (1 + \sqrt{p/n_1})^ 4 - 1$, which decreases as $n_1$ increases.
	The prediction loss of hard parameter sharing for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	  \left|L(\hat{\beta}_2^{\MTL}) - \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right| \le \varepsilon \cdot \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} + p^{-c_{\varphi}} \cdot \OO(d^2 + \sigma^2).
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
%	 {\color{red}[FY: the error also contains $p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) $, both of which cannot be omitted, because (i) there is no assumption on the upper bound of $\kappa^2$, and (ii) we do not necessarily have $c_\varphi<1/4$. We can decide how to present the result concisely (for instance we can impose an upper bound on $\kappa^2$ and that $c_\varphi<1/4$), but it needs to be correct.]}
	 \end{corollary}

@@

The above result provides a more concrete interpretation of the bias-variance decomposition, since the result depends explicitly on datasets' properties of interest.
The proof of Corollary \ref{cor_MTL_loss} can be found in Appendix \ref{app_iso_cov}.
As a remark, by combining the bias and variance limits, we can also obtain a generalization bound for any local minimum of $f(A, B)$.
The proof is similar to Corollary \ref{cor_MTL_loss}, so we omit the details.

Next, we illustrate an interesting phenomenon where adding task one's samples helps task two initially, but hurts eventually.
Consider the limiting estimate in the left hand side of equation \eqref{cor_MTL_error}.
We vary sample ratio $n_1 / n_2$ by fixing $n_2$ and increasing $n_1$.
As a function of the sample ratio, the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ with $n_1$ going to infinity.
We describe a proof sketch.
\begin{enumerate}
	\item The variance term always reduces as $n_1$ increases.
	\item The bias term always increases as $n_1$ increases, which can be verified by calculating the derivative of the bias term.
	\item Moreover, near the point where $n_1$ is zero, one can verify that bias increases less than the decrease in variance, and there is only one positive root for the derivative being zero.
%Combined together, we observe that the \textit{optimal} sample ratio between $n_1$ and $n_2$ should be among one of following three cases:
%(i) $n_1 = 0$, in which case the prediction loss reduces to $\frac{\sigma^2 p}{n_2 - p}$;
%(ii) $n_1 = \infty$, in which case the prediction loss reduces to $d^2$;
%(iii) the critial point where the derivative of the bias and the variance terms w.r.t. $n_1$ equals zero.
\end{enumerate}

%Based on this observation, we have the following trend when we fix $n_2$ and increase $n_1$:
%\begin{itemize}
%	\item If $d^2$ is smaller than $\frac{\sigma^2 p}{n_2 - p}$, then the prediction loss of hard parameter sharing is always lower than STL. That is, task one always transfers positively to task two.
%	\item Otherwise, task one transfers positively to task two initially, but transfers negatively eventually.
%\end{itemize}
%Therefore, while the variance of MTL still reduces compared to STL,
%If the bias increases more than the amount of variance reduced, we observe negative transfer.
%of MTL increases and converges to $pd^2$ as the sample ratio $\rho_1/\rho_2$ increases, and the variance reduction $\frac{\sigma^2}{ \rho_2-1}  -\frac{\sigma^2}{\rho_1+\rho_2-1}$ also increases and converges to $\frac{\sigma^2}{\rho_2-1}$ as $\rho_1/\rho_2$ increases. The tradeoff between these two terms depends on $\rho_1$ in a nonlinear way that is more complicated than the one in the case of varying model distance. But from our results, one can check the following phenomena. If $\frac{\sigma^2}{\rho_2-1} > pd^2$, then the transfer is always positive. On the other hand, if $\frac{\sigma^2}{\rho_2-1} < pd^2$, then there is a transition from the positive transfer to negative transfer as $\rho_1/\rho_2$ increases.
%Therefore, while the variance of MTL still reduces compared to STL,
%If the bias increases more than the amount of variance reduced, we will observe negative transfer.
%Figure \ref{fig_size} provides a simulation result for such a setting (described in Appendix \ref{app_synthetic}).
%We observe that as $n_1 / n_2$ increases, there is a transition from positive to negative transfer.
%There are several unexplained observations near $y = 0$ caused by the concentration error on the right-hand side of \eqref{cor_MTL_error}.

\begin{example}[Covariate shift]\label{ex_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.
Our second example studies the impact of varying covariate shifts on variance in the left hand side of equation \eqref{lem_cov_shift_eq}:
	\[ \bigtr{\Sigma^{(2)} \hat{\Sigma}^{-1}} \rightarrow \frac{1}{n_1 + n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}= \frac{1}{n_1 + n_2} \bigtr{(a_1 M^{\top} M + a_2 \id)^{-1}}. \]
%As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
Hence the variance estimate is determined by the spectrum of $M$.

To illustrate this result, suppose that half of $M$'s singular values are equal to $\lambda > 1$ and the other half are equal to $\lambda^{-1}$.
Hence, when $\lambda = 1$, then there is no covariate shift.
As $\lambda$ increases, the severity of covariate shift increases.
We observe the following dichotomy.
\begin{itemize}
	\item If $n_1 \ge n_2$, then the variance estimate is smallest when there is no covariate shift.
	\item If $n_1 < n_2$, then the variance estimate is largest when there is no covariate shift.
\end{itemize}
\end{example}
To see this, the variance estimate in this example is equal to
\begin{align*}
	\frac{p}{2(n_1 + n_2)}\bigbrace{(\lambda^{-2} a_1 + a_2)^{-1} + (\lambda^2 a_1 + a_2)^{-1}}.
\end{align*}
Denote the above function by $f(\lambda)$ without the scaling $\frac{p}{2(n_1 + n_2)}$.
We compare $f(\lambda)$ to $f(1)$, and observe that
\begin{align*}
	f(\lambda) - f(1) &= {(\lambda^2 a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1})^{-1}
	+ (\lambda^{-2} a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1)^{-1}  - \frac{2(n_1 + n_2)}{n_1 + n_2 - p} \\
	&= (2a_1 - \frac{n_1 + n_2 - p} {n_1 + n_2}) g(\lambda, a_1),
\end{align*}
where $g(\lambda, a_1) > 0$ is a fixed function.
The first step uses equation \eqref{lem_cov_shift_eq} and the second step follows from algebraic calculation (details omitted).
Hence, whether or not $f(\lambda)$ is small than $f(1)$ depends on $a_1$.

We show that $a_1 \ge \frac{(n_1 + n_2 - p)}{2(n_1 + n_2)}$ if and only if $n_1 \ge n_2$.
In Example \ref{ex_covshift}, $a_1$ and $a_2$ have a symmetric role in the self-consistent equations \eqref{lem_cov_shift_eq}.
Hence, when $n_1 \ge n_2$, we have that $a_1 \ge a_2$, hence $a_1 \ge \frac{1}{2}(1 - \frac{p}{n_1 + n_2})$.
The other case is similar.
A formal proof requires solving the self-consistent equations \eqref{lem_cov_shift_eq} and we omit the details.
Thus, we conclude that if $n_1 \ge n_2$, then $f(\lambda) > f(1)$.
If $n_1< n_2$, then $f(\lambda)< f(1)$.
%\begin{proposition}[Source/target sample ratio]\label{prop_data_size}
%	In the isotropic model, suppose that $\rho_1 > 40$ and $\rho_2 > 110$ are fixed constants, and $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
%	Then we have that
%	\begin{itemize}
%		\item \textbf{Positive transfer:} If $\frac{n_1}{n_2} = \frac{\rho_1}{\rho_2} < \frac{1}{\nu} \cdot \frac{1 - 2\rho_2^{-1}}{\Psi(\beta_1, \beta_2) (\rho_2 - 1) - \nu^{-1}}$, then w.h.p. $$\te(\hat{\beta}_2^{\MTL}) < \te(\hat{\beta}_2^{\STL}).$$
%		\item \textbf{Negative transfer:} If $\frac{n_1}{n_2} = \frac{\rho_1}{\rho_2} > {\nu} \cdot \frac{1 - 2\rho_2^{-1}}{\Psi(\beta_1, \beta_2) (\rho_2 - 3/2) - \nu}$, then w.h.p. $$\te(\hat{\beta}_2^{\MTL}) > \te(\hat{\beta}_2^{\STL}).$$
%	\end{itemize}
%\end{proposition}
%Proposition \ref{prop_data_size} describes the bias-variance tradeoff in terms of the sample ratio $\rho_1 / \rho_2$.

%We apply the result to the setting of Figure \ref{fig_size} (described in Appendix \ref{app_synthetic}).
%There are several unexplained observations near $y = 0$ caused by $\nu$.
%The proof of Proposition \ref{prop_data_size} can be found in Appendix \ref{app_proof_32}.

%Proposition \ref{prop_covariate} implies that when $\rho_1\gg \rho_2$, having no covariate shift is the optimal choice for choosing the source task.
%This provides evidence that covariate shift is unfavorable when there are many source task datapoints,
